---
title: "English Premier League 19/20 Performance"
output:
  html_document:
    df_print: paged
---

The data file contains match information from the English premier league during the 19/20 season.

```{r message=FALSE, warning=FALSE}
data <- read.csv("football.csv", header = TRUE)
```

There are 892 matches and 21 features of the data.

```{r}
nrow(data)
ncol(data)
```

The "Score" column contains sometimes invalid scores e.g. "01-Mar". *How should we handle this e.g. discard the rows?*

```{r}
unique(data$Score)
```
There are 132 rows associated with the 20/21 year. *Should these be removed as our questions of interest are for the 19/20 season only?*

```{r}
matches <- data[data$Season == "2020/21",]
nrow(matches)
```

The "Date Range" and "League" features are always the same value. *We can remove these from the data frame.*

```{r}
unique(data$Date.Range)
unique(data$League)
```
The available teams are valid.

```{r}
teams <- unique(data$Team)
awayTeams <- unique(data$AT)
homeTeams <- unique(data$HT)

print(teams)
print(sort(teams) == sort(awayTeams) & sort(teams) == sort(homeTeams))
```

The format of the dates can be DD/MM/YYYY or D/MM/YYYY.

```{r}
head(data$Date, n = 3)
```
The goals scored and conceded values are in the range 0 - 9.

```{r}
unique(data$GS)
unique(data$GA)
```
The points awarded are valid (0, 1, or 3.)

```{r}
unique(data$Pts)
unique(data$xG.Pts)
```
The expected goals are floating point values, but there is no such concept as .1 of a goal.

```{r}
unique(data$xG)
```
The lower and upper bound for expected goals is -0.36 and 0.42 respectively.

```{r}
unique(data$Lower.xG)
unique(data$Upper.xG)
```
There is missing data for Pts vs xG+ and Pts vs xG-. *Both of these can be computed using xG Pts and Pts features.*

```{r}
unique(data$Pts.vs.xG..ve)
unique(data$Pts.vs.xG..ve.1)
```

There is a strong positive relationship between goals scored and expected goals scored for every team.

```{r message=FALSE, warning=FALSE}
cors <- cor(data[,c("GS", "GA", "Pts", "xG", "xGA", "Net.xG", "xG.Pts")])
print(cors)

library(corrplot)
corrplot(cors)
```
The correlation between goals scored and expected goals scored is 0.6, but only when you include observations for all teams.

Breaking it down by team we discover that the relationship ranges from strong to weak depending on the team.

```{r}
team.cors <- data.frame()
for (t in unique(data$Team)) {
  t.data <- subset(data, data$Team == t)
  team.cors <- rbind(team.cors, c(t, cor(t.data$GS, t.data$xG)))
}
colnames(team.cors) <- c("Team", "GS/xG Correlation")
team.cors <- team.cors[order(team.cors$`GS/xG Correlation`, decreasing = TRUE),]
print(team.cors)
```

